---
title: "Computing Group 5 Homework 2"
output: html_notebook
---

Exercises 25-30, 32-36, 38-47

25.

```{r}
logstar <- function (x) {
  
  if (x<1) {0} 
    else {1 + logstar(log(x))}
}

logstar(123)

```

26.

```{r}

gcd <- function (a, b) {
  
  if(a%%b == 0) {b}
  
  else {gcd(b, a%%b)}
  
}

gcd(5, 3)


```

27.
```{r}
seq_is_even <- function(n)
{
  return_vector <- NULL
  for(i in 1:n)
  {
    return_vector=c(rep((i%%2==0),i),return_vector)
  }
  
  return(return_vector)
}

seq_is_even(4)
seq_is_even(3)
seq_is_even(2)
```

28.
```{r}

count_odd <- function(s)
{
  temp_vector <- logical(length(s))
  temp_vector[s%%2==1]=1
  return(sum(temp_vector))
}
count_odd(c(3,3,6,6,7,7))
count_odd(c(1,2,3,4,5,6))

```

29.
```{r}

# We implemented our solution from ex. 26 (GCD) for this
gcd <- function (a, b) {
  
  if(a%%b == 0) {b}
  
  else {gcd(b, a%%b)}
  
}

rel_prime <-function(a,b)
{
  if(gcd(a,b)==1)
  {
    return(TRUE)
  }
  else
  {
    return(FALSE)
  }
}

rel_prime(2,1)
rel_prime(4,2)
rel_prime(2,4)
rel_prime(3,7)


```

30.
```{r}
# We implemented our solution from ex. 26 (GCD) for this
gcd <- function (a, b) {
  
  if(a%%b == 0) {b}
  
  else {gcd(b, a%%b)}
  
}


is_prime <- function(n)
{
 
  for(i in 1:(n-1)){
    if(gcd(n, i)!=1)
    {
      return(FALSE)
    }
  }
  return(TRUE)
}

is_prime(7)
is_prime(3)
is_prime(4)
```



32.
```{r}
proper_divisors = function(x){
  candidates = seq_len(x-1)
  candidates[x%%candidates == 0]
}

perfect_num = function(x) {
  sum_divisors = sum(proper_divisors(x))
  if (sum_divisors == x){
    return(TRUE)
  }
  else {
    return(FALSE)
  }
}

#The first four perfect numbers
for (i in 1:9000){
  if (perfect_num(i)){
    print(i)
  }
}
```


33.
```{r}
divisors_count = function(x){
  candidates = seq_len(x)
  return(length(candidates[x%%candidates == 0]))
}

divisors_count_vec = Vectorize(divisors_count)

highly_composite = function(x) {
  x_divisors = divisors_count(x)
  if (x_divisors > max(divisors_count_vec(1:x-1))){
    return(TRUE)
  }
    else {
      return(FALSE)
    }
}


#The first 20 highly composite numbers
if (highly_composite(1)){
    print(1)
  }

for (i in 2:7561) {
  if (i%%2 != 0){ #odd integers > 1 can't be highly composite, so don't consider them
    next
  }
  
  if (highly_composite(i)){
    print(i)
  }
}

```

34.
```{r}
collatz_step = function(n){
  if (n%%2 == 0){
    n = n/2
  }
  else {
    n = (3*n)+1
  }
}

collatz = function(n_0) {
  n_i = n_0
  while (n_0 != 1) {
    n_0 = collatz_step(n_0)
    n_i = c(n_i, n_0)
  }
  steps = length(n_i) - 1 #exclude n_0, as it does not represent a step
  return(steps)
}


results = sapply(1:100000, collatz)

most_steps = max(results)
most_steps_n = which.max(results)

writeLines(paste("The most steps is for n=", most_steps_n, ", which took", most_steps, "steps"))
```


35.
```{r}
binary_rep = function(x) {
  binary = c()
  n = x
  i = 1
  
  while (n > 0) {
      r = as.logical(n%%2^i)
      binary = c(r, binary)
      n = n - r*2^(i-1)
      i = i+1
  }
  return(as.integer(binary))
}

binary_rep(6)

binary_inv = function(x){
  base = 2^((length(x)-1):0)
  return(sum(x*base))
}

binary_inv(c(1,0,0))


```

36.
Using gcd function from #26
```{r}
lcm = function(a,b) {
  value = abs((a*b))/gcd(a,b)
  return(value)
}

lcm(10,5)
```


38.
```{r}
pascal = function(n){
  line = c(1)
  
  if (n==0){
    return(line)
  }
  
  else {
    for (i in 1:n){
      line = c(0, line) + c(line, 0)
    }
    return(line)
  }
}

for (i in 0:10){
  print(pascal(i))
}
```



39. a.
```{r}
binomial = function(s_0, N, U, D) {
  base = list()
  for (i in 0:N) {
    j = seq(1:(i+1))-1
    base[[i+1]] = s_0 * U^(j) * D^(i-j)
  }
  
  return(base)
}


binomial(1,3,2,0.5)
```

39. b.
```{r}
seq_convert = function(a) {
  return(seq(1:(a+1))-1)
}

binomial_apply = function(s_0, N, U, D) {
  returns = lapply(
    lapply(as.list(seq(1:(N+1))-1), seq_convert), 
    function(i) s_0*U^i*D^(max(i)-i))
  return(returns)
}


binomial_apply(1,3,2,0.5)
```

c.
```{r}
#Efficiency comparison for a small example
start.time = Sys.time()
invisible(binomial(1,3,2,0.5))
end.time = Sys.time()
time.taken = end.time - start.time
time.taken

start.time = Sys.time()
invisible(binomial_apply(1,3,2,0.5))
end.time = Sys.time()
time.taken = end.time - start.time
time.taken


#Efficiency comparison for a large example
start.time = Sys.time()
invisible(binomial(1,3000,2,0.5))
end.time = Sys.time()
time.taken = end.time - start.time
time.taken

start.time = Sys.time()
invisible(binomial_apply(1,3000,2,0.5))
end.time = Sys.time()
time.taken = end.time - start.time
time.taken
```

c. Considering the execution efficiency of each above, the version using apply is typically quicker, though there is some variability between trials. The apply version is a bit easier to read and perhaps reason about to break down the function into its component steps. But, in this example where only one dimension is being iterated over, the 'for' version is fairly straightforward; the difference is probably more pronounced in more complex functions.


40. a.
```{r}
double_sum = function(n) {
  sum = 0
  for (r in 1:n){
    for (c in 1:r) {
      sum = sum + (c^2)/(5+10*r^3)
      }
  }
  return(sum)  
}
double_sum(5)
```

b. 
col returns the column number for each element
```{r}
double_sum_colrow = function (n) {
  base = matrix(0,n,n)
  
  cols = col(matrix(0,n,n))
  cols = apply(cols, c(1,2), function(x) x^2)
  
  rows = row(matrix(0,n,n))
  rows = apply(rows, c(1,2), function(x) 1/(5+10*x^3))
  
  base = rows*cols
  final_sum = sum(base[col(base) <= row(base)]) #sum of lower triangular elements
  return(final_sum)
}

double_sum_colrow(5)
```


41.
```{r}
change_combos = function(a, coin_vec){
  coin_vec = sort(coin_vec, decreasing = TRUE) 
  if (a == 0) {
    return(1)
  }
  else if (a<0) {
    return(0)
  }
  else if (length(coin_vec) == 0) {
    return(0)
  }
  else{
    return(change_combos(a, coin_vec[-1]) + change_combos(a-coin_vec[1], coin_vec))
  }
}

#USD
change_combos(100, c(50, 25, 10, 5, 1))
change_combos(87, c(50, 25, 10, 5, 1))


#Euro
change_combos(100, c(50, 20, 10, 5, 2, 1))
```


42.
```{r}
mod = function(x,y,m) {
  left = (x*y) %% m
  right = (x %% m) * (y %% m)
  
  if (left == right) {
    return(TRUE)
  }
}

mod(10, 5, 3)
mod(15, 5, 3)
mod(1, 4, 3)
mod(1, 3, 4)
mod(100, 2, 4)
mod(10, 5, 4)


expmod=function(a, x, m){
  if (x==0) {
    return(1)
  }
  else if (x%%2==0) {
    return((expmod(a, x/2, m) * (expmod(a, x/2, m)))%%m)
  }
  else {
    return(((a%%m)*expmod(a, x-1, m))%%m)
  }
}

expmod(10,3,6)
(10^3) %% 6

expmod(9,4,2)
(9^4) %% 2

```




