---
title: "QFin Computing Group 2 HW 3"
output: html_notebook
---

48-55, 62-71

62.
```{r}

```


63.
```{r}
a = 1
b = 2^(-53)
c = 2^(-54)

(a+b)+c == a+(b+c)
```

64.
```{r}
log(exp(5)) 
exp(log(5))

log(exp(5)) == exp(log(5)) 

2^(-52) 
log(exp(5)) - exp(log(5))
```
The relationship does hold, "approximately". The difference (representing an error) is the same order of magnitude as the smallest floating point number that double precision computations allow. For most practical purposes, the relationship can be considered true in R.


65.
```{r}
e_limit = function(n) {
  return((1+(1/n))^n)
}

#Estimations for e
e_limit(10^seq(1,20))

#Error terms for all k=1,...,20
abs(e_limit(10^seq(1,20)) - exp(1))
```
The error is minimized at approximately n=10^8. For n larger than that, the error increases and converges to exp(1)-1 at approximately n=10^16. Starting at this n value, the 1/n term becomes smaller than 2^(-52), which is the smallest positive number that R can handle. As a result of this imprecision, the estimate for e has increasingly (and eventually large) error values for n > 10^16.


66.
```{r}
e_approx = function(n, type){
  if (type == "first"){
    value = (1+(1/n))^n
    return(value)
  }
  else if (type == "second") {
    value = (1+n)^(1/n) #note that n=1/h
    return(value)
  }
  else {
    return("Error")
  }
}

exp(1)

for (i in seq(1:25)){
  print(e_approx(10^i, "first"))
}

for (i in seq(1:25)){
  print(e_approx(10^(-i), "second"))
}


#Using either method, e can be approximated to 6 decimal places (2.718282).
```

53. 
```{r}
abs(3* (4/3-1) -1 )

# the idea is that 4/3 gets cut off after 16 decimals. therefore the absolute value results in not being 0, but the smallest possible number. this works in IEEE 754 double precision 
```

54.
```{r}
x <- c(0,7,8)
x[0.9999999999999999]   # the maximum precision we can expect for floating point computations is 16 decimal digits after the comma. 
x[0.99999999999999999]    # therefore r rounds this to 1
```

55.
```{r}
options(digits=22)

ex55a <- function(x) {
  x <- 1
  while(x + 1 != 1)
    x <- x/2
  return(x*2)
}

ex55a()

1+ex55a()

ex55b <- function(x) {
  x <- 1
  while(1 - x != 1)
    x <- x/2
  return(x*2)
}

ex55b()
1-ex55b()

```
